---
title: "2021 01 28 Exercises"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(keras)
use_condaenv("r-reticulate")
```

### Parkinsons

```{r}
pdata <- read_csv("input/parkinsons_updrs.data") %>%
  rename(subjectID=`subject#`)
head(pdata)
```


set up test and train, by subject

```{r}
length(unique(pdata$subjectID))
unique(pdata$subjectID)
```

Let's set 6 subjects aside as the test set.  It would be nice to do a cross validation by subjID but for now I am realy short on time, so just do one validation.

```{r}
set.seed(3434)
testsubj <- sample(1:42, size=6)

motor_UPDRS.test <- pdata %>% filter(subjectID %in% testsubj) %>% pull(motor_UPDRS)
motor_UPDRS.train <- pdata %>% filter(! subjectID %in% testsubj) %>% pull(motor_UPDRS)

pdata.test <- pdata %>% filter(subjectID %in% testsubj) %>% select(-test_time, -motor_UPDRS, -total_UPDRS)
pdata.train <-  pdata %>% filter(! subjectID %in% testsubj) %>% select(-test_time, -motor_UPDRS, -total_UPDRS)

valsubj <- sample(unique(pdata.train$subjectID), size=6)

pdata.train.partial <- pdata.train %>% filter(! subjectID %in% valsubj)
pdata.train.val <- pdata.train %>% filter(subjectID %in% valsubj)

motor_UPDRS.train.partial <- motor_UPDRS.train[! pdata.train$subjectID %in% valsubj]
motor_UPDRS.val <- motor_UPDRS.train[pdata.train$subjectID %in% valsubj]

trainID <- pdata.train %>% pull(subjectID)
train.partialID <- pdata.train.partial %>% pull(subjectID)
valID <- pdata.train.val %>% pull(subjectID)
testID <- pdata.test %>% pull(subjectID)

pdata.train <- pdata.train %>% select(-subjectID)
pdata.train.partial <- pdata.train.partial %>% select(-subjectID)
pdata.train.val <- pdata.train.val %>% select(-subjectID)
pdata.test <- pdata.test %>% select(-subjectID)

```

center and scale the predictors

```{r}
(pcenter <- pdata.train %>% apply(2, mean))
(psd <- pdata.train %>% apply(2, sd))

pdata.train <- scale(pdata.train, center=pcenter, scale = psd)
pdata.test <- scale(pdata.test, center=pcenter, scale = psd)
pdata.train.partial <- scale(pdata.train.partial, center=pcenter, scale = psd)
pdata.train.val <- scale(pdata.train.val, center=pcenter, scale = psd)
```

build and compile model, fit
```{r}
pmodel <- keras_model_sequential() %>%
  layer_dense(units = 16, input_shape=ncol(pdata.train), activation = "relu") %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1)

pmodel <- pmodel %>% compile(
  optimizer = "rmsprop", 
    loss = "mse", 
    metrics = c("mae")
  )

system.time(phistory <- pmodel %>% fit(
    pdata.train.partial, 
    motor_UPDRS.train.partial,
    validation_data = list(pdata.train.val, 
                           motor_UPDRS.val),
    epochs = 20, 
    batch_size = 128, 
    verbose = 0
))
```

```{r}
plot(phistory)
```

refit with full train set, 6 epochs

```{r}
pmodel <- keras_model_sequential() %>%
  layer_dense(units = 16, input_shape=ncol(pdata.train), activation = "relu") %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1)

pmodel <- pmodel %>% compile(
  optimizer = "rmsprop", 
    loss = "mse", 
    metrics = c("mae")
  )

system.time(pmodel %>% fit(
    pdata.train, 
    motor_UPDRS.train,
    epochs = 6, 
    batch_size = 128, 
    verbose = 0
))
```

```{r}
presults <- pmodel %>% evaluate(pdata.test, motor_UPDRS.test)
presults
```

how bad is this?  what is range of motor_UPDRS?

```{r}
ppredictions <- pmodel %>% predict(x=pdata.test)
```


```{r}
qplot(x=motor_UPDRS.test, y=ppredictions, color=as.factor(testID))
```

```{r}
ppredictions.train <- pmodel %>% predict(x=pdata.train)
```


```{r}
qplot(x=motor_UPDRS.train, y=ppredictions.train)
```

probably hopeless

## Abalone

```{r}
adata <- read_csv("input/abalone.data", 
                  col_names = c("sex",
                                "length",
                                "diameter",
                                "height",
                                "wholeWeight",
                                "shuckedWeight",
                                "visceraWeight",
                                "shellWeight",
                                "rings"))
head(adata)
dim(adata)
```

format it, etc
```{r}
atarget <- adata %>% pull(rings)
adata <- adata %>% mutate(sex=as.integer(as.factor(sex))) %>%
  select(-rings)

train <- sample(nrow(adata), size=3200)
val <- sample(train, size=320)

adata.train <- adata[train,]
adata.test <- adata[-train,]
adata.val <- adata[val,]
adata.train.part <- adata[setdiff(train, val),]

atarget.train <- atarget[train]
atarget.test <- atarget[-train]
atarget.val <- atarget[val]
atarget.train.part <- atarget[setdiff(train, val)]
```

scale and center
```{r}
acenter <- apply(adata.train,2, mean)
asd <- apply(adata.train,2, sd)

adata.train <- scale(adata.train, center=acenter, scale=asd)
adata.test <- scale(adata.test, center=acenter, scale=asd)
adata.train.part <- scale(adata.train.part,center=acenter, scale=asd)
adata.val <- scale(adata.val, center=acenter,scale=asd)
```


## regression problem.

setup model
```{r}
amodel <- keras_model_sequential() %>%
  layer_dense(units=8, activation="relu", input_shape = ncol(adata.train)) %>%
  layer_dense(units=8, activation="relu") %>%
  layer_dense(units=1)

amodel <- amodel %>% compile(
  optimizer = "rmsprop", 
    loss = "mse", 
    metrics = c("mae")
  )

system.time(ahistory <- amodel %>% fit(
    adata.train.part, 
    atarget.train.part,
    validation_data=list(adata.val, atarget.val),
    epochs = 100, 
    batch_size = 128, 
    verbose = 0
))
```

```{r}
plot(ahistory)
```

25 epochs looks good
```{r}
amodel <- keras_model_sequential() %>%
  layer_dense(units=8, activation="relu", input_shape = ncol(adata.train)) %>%
  layer_dense(units=8, activation="relu") %>%
  layer_dense(units=1)

amodel <- amodel %>% compile(
  optimizer = "rmsprop", 
    loss = "mse", 
    metrics = c("mae")
  )

system.time(ahistory <- amodel %>% fit(
    adata.train, 
    atarget.train,
    epochs = 25, 
    batch_size = 128, 
    verbose = 0
))
```

```{r}
amodel %>% evaluate(adata.test, atarget.test)
```

```{r}
apredict <- amodel %>% predict(x=adata.test)
```

```{r}
qplot(atarget.test, apredict, alpha=0.5)
```

```{r}
cor(apredict, atarget.test)
```

```{r}
range(apredict)
range(atarget.test)
apredict.int <- round(apredict)
```

```{r}
sum(apredict.int==atarget.test)/length(apredict.int)
```


