---
title: "7.2_TryingTensorBoard"
author: "Julin Maloof"
date: "11/21/2021"
output:
  html_document:
    theme: cerulean
    highlight: testmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(keras)
use_condaenv("r-reticulate")
```


This is mostly from section 6.4, modified for 7.2

## Implementing a 1D convnet
  
In Keras, you use a 1D convnet via the `layer_conv_1d()` function, which has an interface similar to `layer_conv_2d()`. It takes as input 3D tensors with shape `(samples, time, features)` and returns similarly shaped 3D tensors. The convolution window is a 1D window on the temporal axis: the second axis in the input tensor.

Let's build a simple two-layer 1D convnet and apply it to the IMDB sentiment-classification task you're already familiar with. As a reminder, this is the code for obtaining and preprocessing the data.

```{r}
max_features <- 2000
max_len <- 500

cat("Loading data...\n")
imdb <- dataset_imdb(num_words = max_features)
c(c(x_train, y_train), c(x_test, y_test)) %<-% imdb 
cat(length(x_train), "train sequences\n")
cat(length(x_test), "test sequences")

cat("Pad sequences (samples x time)\n")
x_train <- pad_sequences(x_train, maxlen = max_len)
x_test <- pad_sequences(x_test, maxlen = max_len)
cat("x_train shape:", dim(x_train), "\n")
cat("x_test shape:", dim(x_test), "\n")
```

1D convnets are structured in the same way as their 2D counterparts, which you used in chapter 5: they consist of a stack of `layer_conv_1d()` and `layer_max_pooling_1d()`, ending in either a global pooling layer or a `layer_flatten()`, that turn the 3D outputs into 2D outputs, allowing you to add one or more dense layers to the model for classification or regression.

One difference, though, is the fact that you can afford to use larger convolution windows with 1D convnets. With a 2D convolution layer, a 3 × 3 convolution window contains 3 * 3 = 9 feature vectors; but with a 1D convolution layer, a convolution window of size 3 contains only 3 feature vectors. You can thus easily afford 1D convolution windows of size 7 or 9.

This is the example 1D convnet for the IMDB dataset.

```{r}
model <- keras_model_sequential() %>% 
  layer_embedding(input_dim = max_features, output_dim = 128,
                  input_length = max_len) %>% 
  layer_conv_1d(filters = 32, kernel_size = 7, activation = "relu") %>% 
  layer_max_pooling_1d(pool_size = 5) %>% 
  layer_conv_1d(filters = 32, kernel_size = 7, activation = "relu") %>% 
  layer_global_max_pooling_1d() %>% 
  layer_dense(units = 1)

summary(model)

model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "binary_crossentropy",
  metrics = c("acc")
)
```


```{r}
if(!dir.exists("7.2_log.dir")) dir.create("7.2_log.dir")

tensorboard("7.2_log.dir")

callbacks <- list(
  callback_tensorboard(
    log_dir = "7.2_log.dir",
    histogram_freq = 1
  )
)

history <- model %>% fit(
  x_train, y_train,
  epochs = 20,
  batch_size = 128,
  validation_split = 0.2,
  callbacks = callbacks
)
```


```{r}
plot(history)
```
