---
title: "Chapter 6.3"
output: html_notebook
---

```{r}
library(keras)
use_condaenv("r-reticulate")
```


```{r, eval=FALSE, echo=TRUE}
dir.create("jena_climate", recursive = TRUE)
download.file(
  "https://s3.amazonaws.com/keras-datasets/jena_climate_2009_2016.csv.zip",
  "jena_climate/jena_climate_2009_2016.csv.zip"
)
unzip(
  "jena_climate/jena_climate_2009_2016.csv.zip",
  exdir = "jena_climate"
)
```


Let's take a look at the data:

```{r}
library(tibble)
library(readr)

data_dir <- "jena_climate"
fname <- file.path(data_dir, "jena_climate_2009_2016.csv")
data <- read_csv(fname)
```

```{r}
glimpse(data)
```

Here is the plot of temperature (in degrees Celsius) over time. On this plot, you can clearly see the yearly periodicity of temperature.

```{r}
library(ggplot2)
ggplot(data, aes(x = 1:nrow(data), y = `T (degC)`)) + geom_line()
```

On this plot, you can clearly see the yearly periodicity of temperature.

Here is a more narrow plot of the first ten days of temperature data (since the data is recorded every ten minutes, we get 144 data points per day):

```{r}
ggplot(data[1:1440,], aes(x = 1:1440, y = `T (degC)`)) + geom_line()
```

On this plot, you can see daily periodicity, especially evident for the last 4 days. We can also note that this ten-days period must be coming from a fairly cold winter month.

If we were trying to predict average temperature for the next month given a few month of past data, the problem would be easy, due to the reliable year-scale periodicity of the data. But looking at the data over a scale of days, the temperature looks a lot more chaotic. So is this timeseries predictable at a daily scale? Let's find out.

```{r}
sequence_generator <- function(start) {
  value <- start - 1
  function() {
    value <<- value + 1
    value
  }
}

gen <- sequence_generator(10)
gen()
gen()
gen
```

## Preprocess data

```{r}
data <- data.matrix(data[,-1])
```

scale it
```{r}
train_data <- data[1:200000,]
mean <- apply(train_data, 2, mean)
std <- apply(train_data, 2, sd)
data <- scale(data, center = mean, scale = std)
```

## make the generator

```{r}
generator <- function(data, lookback, delay, min_index, max_index,
                      shuffle = FALSE, batch_size = 128, step = 6) {
  if (is.null(max_index))
    max_index <- nrow(data) - delay - 1
  i <- min_index + lookback
  function() {
    if (shuffle) {
      rows <- sample(c((min_index+lookback):max_index), size = batch_size)
    } else {
      if (i + batch_size >= max_index)
        i <<- min_index + lookback
      rows <- c(i:min(i+batch_size, max_index))
      i <<- i + length(rows)
    }
    
    samples <- array(0, dim = c(length(rows), 
                                lookback / step,
                                dim(data)[[-1]]))
    targets <- array(0, dim = c(length(rows)))
                     
    for (j in 1:length(rows)) {
      indices <- seq(rows[[j]] - lookback, rows[[j]], 
                     length.out = dim(samples)[[2]])
      samples[j,,] <- data[indices,]
      targets[[j]] <- data[rows[[j]] + delay,2]
    }            
    
    list(samples, targets)
  }
}
```


```{r}
lookback <- 1440
step <- 6
delay <- 144
batch_size <- 128

train_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 1,
  max_index = 200000,
  shuffle = TRUE,
  step = step, 
  batch_size = batch_size
)

val_gen = generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 200001,
  max_index = 300000,
  step = step,
  batch_size = batch_size
)

test_gen <- generator(
  data,
  lookback = lookback,
  delay = delay,
  min_index = 300001,
  max_index = NULL,
  step = step,
  batch_size = batch_size
)

# This is how many steps to draw from `val_gen`
# in order to see the whole validation set:
val_steps <- (300000 - 200001 - lookback) / batch_size

  # This is how many steps to draw from `test_gen`
# in order to see the whole test set:
test_steps <- (nrow(data) - 300001 - lookback) / batch_size
```

## baseline

```{r}

evaluate_naive_method <- function() {
  batch_maes <- c()
  for (step in 1:val_steps) {
    c(samples, targets) %<-% val_gen()
    preds <- samples[,dim(samples)[[2]],2]
    mae <- mean(abs(preds - targets))
    batch_maes <- c(batch_maes, mae)
  }
  print(mean(batch_maes))
}

evaluate_naive_method()
```
note: this is in standardized units

## Simple cNN

```{r, echo=TRUE, results='hide'}

model <- keras_model_sequential() %>% 
  layer_flatten(input_shape = c(lookback / step, dim(data)[-1])) %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1)

model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)

history <- model %>% fit_generator(
  train_gen,
  steps_per_epoch = 500,
  epochs = 20,
  validation_data = val_gen,
  validation_steps = val_steps
)
```

Let's display the loss curves for validation and training:

```{r}
plot(history)
```

## First recurrent

```{r echo=TRUE, results='hide'}
model <- keras_model_sequential() %>% 
  layer_gru(units = 32, input_shape = list(NULL, dim(data)[[-1]])) %>% 
  layer_dense(units = 1)

model %>% compile(
  optimizer = optimizer_rmsprop(),
  loss = "mae"
)

history <- model %>% fit_generator(
  train_gen,
  steps_per_epoch = 500,
  epochs = 20,
  validation_data = val_gen,
  validation_steps = val_steps
)
```

Let look at our results:

```{r}
plot(history)
```



