---
title: "ChickenPox"
author: "Julin Maloof"
date: "7/12/2021"
output: html_document
---

```{r}
library(tidyverse)
```


The data consists of weekly chickenpox incident rates for every county in Hungary from 2005 to 2015

The goal is to be able to predict chickenpox disease rates in the future

The data set is at [UCI repository](https://archive.ics.uci.edu/ml/datasets/Hungarian+Chickenpox+Cases)

Note that there are two data files.  `hungary_chickenpox.csv` has case counts per county.  `hungary_county_edges.csv` is an adjacency matrix indicating which counties abut one another, but I am not sure how to include this.

Exercises:

0. Do you want to center and scale?  If so don't forget to do it.

1. plot the data in an informative way

2. Because this is a small data set and we are just practicing we are not going to set up a test set, just training and validation.  Set up your generators or data so that training data is 2005-2012 and validation is 2013, 2014.

3. Predict chickenpox case # in Budapest 2 weeks in the future.  Be sure to include a baseline model for comparison.  

4. Recode the data so that you can make predicition for all counties 2 weeks in the future.  Be sure to include a baseline model for comparison.

5. If you are getting good predictions, how good can you do at 1 month?  If predictions are poor maybe try going down to 1 week.

Get the data
```{r}
if(!dir.exists("chickenpox")) {
  dir.create("chickenpox")
  download.file(url="https://archive.ics.uci.edu/ml/machine-learning-databases/00580/hungary_chickenpox.zip",
  destfile="chickenpox/hungary_chickenpox.zip")
  unzip("chickenpox/hungary_chickenpox.zip", exdir = "chickenpox")
} 
```

plot it
```{r}
pox <- read_csv("chickenpox/hungary_chickenpox.csv")
```

```{r}
head(pox)
```

```{r}
pox <- pox %>% 
  mutate(Date = lubridate::dmy(Date),
         year = lubridate::year(Date),
         week=lubridate::week(Date)) %>%
  select(Date, week, year, everything())
```

```{r}
pox %>% pivot_longer(-c(Date, week, year), names_to = "county") %>%
  ggplot(aes(x=week, y=value, color=county)) +
  geom_line() +
  facet_wrap(~ year)
  
```

scale and center.  Do each county separately.

```{r}
countymeans <- pox %>% 
  filter(year < 2013) %>% 
  summarize(across(.cols=c(-Date, -week, -year), mean)) %>%
  as.numeric

countysd <- pox %>% 
    filter(year < 2013) %>% 
  summarize(across(.cols=c(-Date, -week, -year), sd)) %>%
  as.numeric()

poxscale <- pox
poxscale[,-1:-3] <- scale(pox[,-1:-3], center = countymeans, scale = countysd)
```

did it work?
```{r}
poxscale %>% 
  filter(year < 2013) %>% 
  summarize(across(.cols=c(-Date, -week, -year), mean)) 

poxscale %>% 
    filter(year < 2013) %>% 
  summarize(across(.cols=c(-Date, -week, -year), sd))
```
yes


Reformat for exercise 4:

```{r}
poxl <- poxscale %>% 
  pivot_longer(-c(Date, week, year), names_to = "county", values_to = "cases") %>%
  pivot_wider(id_cols = -county, names_from = county, values_from = county, values_fn = function(x) 1, values_fill = 0 )
poxl 
```

